A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-11 ISMB/methods_ISMB.tex
Added highlight macro and fixed response to reviewers (still to be completed)

\subsubsection{Precomputing the \GCWeighted partition function}\label{sec:pf}

Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function ($x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$). 
Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
\begin{equation}
	\Z{\BoolTrue,\varepsilon}{a,b}=\Z{\BoolFalse,\varepsilon}{a,b}= e^{-0/RT} = 1.
	\label{eq:Z_in}
\end{equation}

@@

Remark that the number of combinations of $a$, $b$ and $\N$ remains bounded by a constant, thus the complexity of computing $\Z{\N,\Struct}{a,b}$ mainly depends on the values taken by $\Struct$ upon subsequent recursive calls. Such values are entirely determined by $\Struct$ at any given step of the recursion, and their dependency can be summarized in a tree having $\Theta(|\Struct|)$. Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.

@@

\subsubsection{Self-adaptive sampling strategy}

@@

It was previously established~\citep{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, the distribution of \GCContent is highly concentrated, as asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp. $\kappa\in\Omega(1/\sqrt n)$) grows like $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. a constant), leading to the announced complexities. Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0$, i.e. without any tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).

>>>>

\highlight{
\subsubsection{Precomputing the \GCWeighted partition function}\label{sec:pf}

Firstly, a dynamic programming algorithm computes $\Z{\N,\Struct}{a,b}$ the \GCWeighted partition function (the dependency in $x$ is omitted here for the sake of clarity) for a structure $\Struct$, assuming its (previously chosen) flanking nucleotides are $a$ and $b$ respectively, either forming a closing base-pair ($\N=\BoolTrue$) or not ($\N=\BoolFalse$).
Remark that the empty structure only supports the empty sequence, having energy $0$, so one has
\begin{equation}
	\Z{\BoolTrue,\varepsilon}{a,b}=\Z{\BoolFalse,\varepsilon}{a,b}= e^{-0/RT} = 1.
	\label{eq:Z_in}
\end{equation}

@@

Remark that the number of combinations of $a$, $b$ and $\N$ remains bounded by a constant, thus the complexity of computing $\Z{\N,\Struct}{a,b}$ mainly depends on the values taken by $\Struct$ upon subsequent recursive calls. Such values are entirely determined by $\Struct$ at any given step of the recursion, and their dependency can be summarized in a tree having $\Theta(|\Struct|)$. Therefore, the computation of $\Z{\N,\Target}{a,b}$ requires $\Theta(n)$ time and space using dynamic-programming.}

@@

\highlight{
\subsubsection{Self-adaptive sampling strategy}

@@

It was previously established~\citep{Waldispuhl2011} that, for each value of $x$, there exists constants $\mu_x$ and $\sigma_x$ such that the distribution of \GCContent asymptotically converges towards a normal law having expectation in $\mu_x\cdot n\cdot(1+o(1))$ and standard deviation in $\sigma_x\cdot\sqrt{n}\cdot(1+o(1))$.
Furthermore, the distribution of \GCContent is highly concentrated, as asserted by its limited standard deviation, therefore the expected number of attempts required to generate a valid sequence when $\kappa=0$ (resp. $\kappa\in\Omega(1/\sqrt n)$) grows like $\Theta(\sqrt{n})$ (resp. $\Theta(1)$, i.e. a constant), leading to the announced complexities. Formally, since a suitable weight $x$ must be recomputed for each targeted structure and \GCContent, then the number $M$ of iterations required for the converge can be accounted for explicitly, leading to time complexities in $\Theta((M+\sqrt{n})\cdot k\cdot n)$ (if $\kappa=0$, i.e. without any tolerance) and $\Theta(M\cdot k\cdot n)$ (if $\kappa>0$).
}
